---
title: "Breast Cancer Diagnosis Classification Analysis"
author: Austin Lackey, Ethan Powers and Danny Laposata
date: "December 9th, 2022"
output: 
  powerpoint_presentation: 
    reference_doc: template.pptx
  
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Background

- Breast Cancer Diagnosis Data
- Predict Diagnosis of Tumor based on features
- Reduce Type II Error (False-Negatives)
- Feature Selection

## Data

- Data collected from Kaggle
- Response: Diagnosis (Benign/Malignant)
- Mean, SE, and "Worst" measurements of 10 variables
- Training/Test Split (60/40)

## Features

- 
- 

## EDA

Different measurements of each variable are correlated

Relationship with Diagnosis visible for each measurement

```{r, fig.height=6, fig.width=8}
source("PPTX-code.R", echo = F, local = T)


```




## Classification Analysis: Austin

## Classification Analysis: Danny

- LogReg
- KNN
- QDA
- RF
- xGBoost
- SVC
- Lower prediction cutoff to account for Type II Error

## K-Nearest Neigbors Model

- Tuned neighbors value
- 2 False-Negatives
- 96.04% prediction accuracy

## Support Vector Classifier Model

- Tuned cost value
- 2 False-Negatives
- 96.48% prediction accuracy

## Random Forest Model

- 1 False-Negative
- 95.15% prediction accuracy

## Regression Analysis: Ethan



## Model Assesments





## Works Cited

- Insert link to kaggle dataset here