---
title: "Predicting Covid-19 Data"
author: "Austin Lackey, Chien Lin and Jin Peng"
date: "Tuesday, December 6th, 2022"
output:
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# useful libraries
library(tidyverse)
library(pracma)
library(minpack.lm)
# read covid data
data <- read_csv("owid-covid-data.csv")
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```


## The Monte Carlo Study: 

**Mathematical prediction of the time evolution of the COVID-19 pandemic in Italy by a Gauss error function and Monte Carlo simulations**


<br>

*Author: Ignazio Ciufolini, Antonio Paolozzi*


## Introduction : Ciufolini and Paolozzi's Study

* The time evolution of the number of cumulative diagnosed positive cases and fatalities in China approximates the distribution of a gauss error function: $a+b*erf(c*x-d)$.
* Performed the function on Italy data to approximate the number of cases.
* Performed 150 Monte Carlo simulations to predict the day in which the peak of the number of daily cases in Italy occurs.

# China Simulation

## Available Official Data (Positive Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
# extract China data of interest
china <- data |>
  filter(location == "China") |>
  filter(date >= "2020-01-22", date <= "2020-03-28") |>
  select(c(date, total_cases, total_deaths)) |>
  mutate(days = 1:n())

# recreate plots
positive <- ggplot(data = china) +
  geom_point(aes(x = date, y = total_cases)) +
  labs(x = "Date", y = "China cumulative number of cases")
positive
```

## Available Official Data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
ggplot(data = china) +
  geom_point(aes(x = date, y = total_deaths)) +
  labs(x = "Date", y = "China cumulative number of fatalities") +
  ylim(0,5000)
```

## Fit by the Gauss Error Function

Apply an $nls$ function to determine the nonlinear (weighted) least-squares estimates of the four parameters $a,b,c,d$ of a Gauss error function model: $a+b*erf(c*x-d)$.

```{r, include=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
```

```{r, message=FALSE, warning=FALSE}
model <- nls(total_cases ~ gauss_fn(days,a,b,c,d), 
             data=china, 
             start=list(a=10000,b=10000,c=0.1,d=1))
```

## Fit by the Gauss Error Function and Peak Day (Positive Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
try <- china$total_cases
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=seq(1,67,length.out=6700)
tmp <- eval(fn_try_deriv)
peak1 <- min(x[tmp < 0])
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
model1 <- nls(total_cases ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
# summary(model1)$parameters
fn_model1 <- function(x) {4.017e+04 + 4.047e+04 * erf(9.284e-02 * x - 1.732e+00)}
ggplot(data = china) +
  geom_point(aes(x = days, y = total_cases)) +
  labs(x = "Days", y = "China cumulative number of cases") +
  stat_function(fun=fn_model1, color="#d33636", size=1.5) +
  geom_vline(xintercept = peak1, color="#d33636", size=1.5, linetype="dashed")
```

## Fit by the Gauss Error Function and Peak Day (Fatalities)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
try <- china$total_deaths
model_try <- nls(try ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=2200,b=2200,c=0.1,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

# find the second derivative of fitted function
# source: https://proofwiki.org/wiki/Derivative_of_Error_Function
erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}
fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=seq(1,67,length.out=6700)
tmp <- eval(fn_try_deriv)
peak2 <- min(x[tmp < 0])
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
model2 <- nls(total_deaths ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
#summary(model2)$parameters
fn_model2 <- function(x) {1.571e+03 + 1.658e+03 * erf(6.194e-02 * x - 1.533e+00)}
ggplot(data = china) +
  geom_point(aes(x = days, y = total_deaths)) +
  labs(x = "Days", y = "China cumulative number of fatalities") +
  ylim(0,5000) +
  stat_function(fun=fn_model2, color="#1b95ff", size=1.5) +
  geom_vline(xintercept = peak2, color="#1b95ff", size=1.5, linetype="dashed")
```

# Italy Simulation

## Initial Data for Cases in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(400)
# extract Italy data
italy <- data %>% 
  filter(location=="Italy") %>% 
  filter(date <= "2020-04-21") %>%
  filter(date >= "2020-02-15") %>%
  select(date, total_cases, total_deaths)
  #ref: %>% mutate(days = c(1:67))
italy[45:67,] <- NA
italy <- italy %>% mutate(days = c(1:67))

#plot
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")
```

## Initial Data for Fatalities in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")
```

## Fitting Gauss Error Function to Positive Cases in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- 1:67

c <- summary(nlsLM(total_cases~gauss_fn(days,a,b,c,d), data=italy, start=list(a=50000,b=50000,c=.01,d=1)))$parameters
d <- summary(nlsLM(total_deaths~gauss_fn(days,a,b,c,d), data=italy, start=list(a=10000,b=10000,c=.01,d=1)))$parameters

cf_italy <- function(x) c[1]+c[2]*erf(c[3]*x-c[4])
df_italy <- function(x) d[1]+d[2]*erf(d[3]*x-d[4])

#plot
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")
```

## Fitting Gauss Error Function to Fatalities in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=df_italy, color="#1b95ff", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")
```

## Overview of the Monte Carlo Simulation

* Create a random matrix : m X n
  * m = number of random outcomes (*chosen to be 150*)
  * n = number of observed days (*n = 1,...,j*)
* Each # in the matrix is a Gaussian distribution with mean = 1 and sigma = 0.1
  * Multiply each column (*j*) by the # of total cases (or deaths) that corresponds to each day *j*
* Each of the 150 simulations were fitted with the Gauss error function
* Then determined the date of the flex for each simulation

## Using MonteCarlo Simulation to predict the day of flex

- The mean of the day of flex for cases is 40.59 days and the standard deviation is 5.222832 days.
- The mean of the day of flex for fatalities is 43.46 days and the standard deviation is 5.228381 days.

## MonteCarlo Simulation Data (Cases) (150 trials)

Each red dot corresponds to the day of occurrence of the flex obtained with each of the 150 Monte Carlo simulations

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
n <- 44
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

italy_mod <- italy[1:44,]

model_ca <- c()
for (i in 1:n) {
    model_ca <- c(model_ca, italy_mod$total_cases[i]*gau_sample[,i])
}
model_ca <- matrix(model_ca, nrow = m, ncol = n)

model_fa <- c()
for (i in 1:n) {
    model_fa <- c(model_fa, italy_mod$total_deaths[i]*gau_sample[,i])
}
model_fa <- matrix(model_fa, nrow = m, ncol = n)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# cases
flex_ca <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_ca[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=50000,b=50000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_ca[i] <- min(x[tmp < 0])
}
mean(flex_ca)
sd(flex_ca)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
    geom_point(aes(x=1:m, y=flex_ca), color="#d33636") +
    geom_hline(yintercept=mean(flex_ca), color="#d33636", linetype="dashed", size=1.5) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Cases)", x = "Trial", y = "Day of Flex")
```

## Applying the day of flex indicator to the plot (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot(cs)
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases") + 
  geom_vline(xintercept = mean(flex_ca), color="#d33636", size=1, linetype="dashed")
```

## MonteCarlo Simulation Data (Fatalities) (150 trials)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# fatalities
flex_fa <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_fa[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=15000,b=15000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_fa[i] <- min(x[tmp < 0])
}
flex_fa <- flex_fa[-c(12,112,124,133)]
mean(flex_fa)
sd(flex_fa)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
    geom_point(aes(x=1:146, y=flex_fa), color="#1b95ff") +
    geom_hline(yintercept=mean(flex_fa), color="#1b95ff", linetype="dashed", size=1.5) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Fatalities)", x = "Trial", y = "Day of Flex")
```

## Applying the day of flex indicator to the plot (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot(ft)
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() +
    stat_function(fun=df_italy, color="#1b95ff", size=1.5, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths") +
  geom_vline(xintercept = mean(flex_fa), color="#1b95ff", size=1, linetype="dashed")
```


# Australia Simulation

## Initial Data for Australia (Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
australia <- data %>% filter(location == "Australia", date > "2020-06-01", date < "2020-12-19")
australia <- subset(australia, select = c(date, total_cases, total_deaths))
australia <- australia %>% mutate(days = 1:n())
minimizedDays <- 70
australiaPre <- australia %>% filter(days < minimizedDays)
cases_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_Australia)
```

## Initial Data for Australia (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_Australia)
```

## Fitting a Gauss Error Function to Australia Data (Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
x <- 0:minimizedDays


gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
cf <- summary(nlsLM(total_cases ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=10000,b=500,c=.02,d=1)))$parameters
ff <- summary(nlsLM(total_deaths ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=100,b=100,c=.01,d=1)))$parameters


cases_function_australia <- function(x) cf[1]+cf[2]*erf(cf[3]*x-cf[4])
fatalities_function_australia <- function(x) ff[1]+ff[2]*erf(ff[3]*x-ff[4])


cases_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_pre_plot_Australia)
```

## Fitting a Gauss Error Function to Australia Data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_pre_plot_Australia)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
x <- 0:200

# cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
# fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)

cases_plot_Australia <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Cases in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    labs(title = "Total Deaths in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")
```

## Applying the function to the future dataset (Cases)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
cases_plot_Australia_future <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    xlim(0, 200) +
    labs(title = "Extrapolating Future Case Data with Gauss Error Function", x = "Days", y = "Total Cases")
fatalities_plot_Australia_future <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    xlim(0, 200) +
    labs(title = "Extrapolating Future Fatality Data with Gauss Error Function", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia_future)
```

## Applying the function to the future dataset (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia_future)
```

# MonteCarlo Simulation for Australia's Day of Flex

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Montecarlo Simulation
set.seed(400)
n <- dim(australia)[1]
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

mc_data_ca <- c()
for (i in 1:n) {
    mc_data_ca <- c(mc_data_ca, australia$total_cases[i]*gau_sample[,i])
}
mc_data_ca <- matrix(mc_data_ca, nrow = m, ncol = n)
mc_data_fa <- c()
for (i in 1:n) {
    mc_data_fa <- c(mc_data_fa, australia$total_deaths[i]*gau_sample[,i])
}
mc_data_fa <- matrix(mc_data_fa, nrow = m, ncol = n)

day_of_flex_ca <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_ca[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=15000,b=9000,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_ca[i] <- min(x[temp < 0])
}
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
day_of_flex_fa <- rep(NA, m)
for (i in 1:m) {
    trial <- mc_data_fa[i,]
    trial_model <- nlsLM(trial ~ gauss_fn(days,a,b,c,d), data=australia, start=list(a=300,b=200,c=.01,d=1))
    model_coef <- summary(trial_model)$parameters[, 1]
    # trial_function <- function(x, coef) {coef[1] + coef[2]*erf(coef[3]*x-coef[4])}
    gauserr_derivative <- {2/sqrt(pi)*exp(-x^2)}
    function_trial_derivative <- D(bquote(.(model_coef[2])*2/sqrt(pi)*exp(-(.(model_coef[3])*x-.(model_coef[4]))^2)), 'x')
    x = seq(1, 200, length.out=6700)
    temp <- eval(function_trial_derivative)
    day_of_flex_fa[i] <- min(x[temp < 0])
}
```

## MonteCarlo Simulation Data (Cases) (150 trials)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
australiaMCPlot_cases <- ggplot() +
    geom_point(aes(x=1:m, y=day_of_flex_ca), color="#d33636") +
    geom_hline(yintercept=mean(day_of_flex_ca), color="#d33636", linetype="dashed", size=2) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Cases)", x = "Trial", y = "Day of Flex")
australiaMCPlot_fatalities <- ggplot() +
    geom_point(aes(x=1:m, y=day_of_flex_fa), color="#1b95ff") +
    geom_hline(yintercept=mean(day_of_flex_fa), color="#1b95ff", linetype="dashed", size=2) + 
    labs(title = "MonteCarlo Simulation for Day of Flex (Fatalities)", x = "Trial", y = "Day of Flex")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(australiaMCPlot_cases)
```

## MonteCarlo Simulation Data (Fatalities) (150 trials)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(australiaMCPlot_fatalities)
```

## Using MonteCarlo Simulation to Predict the Day of Flex

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
print(mean(day_of_flex_fa))
print(sd(day_of_flex_fa))
print(mean(day_of_flex_ca))
print(sd(day_of_flex_ca))
```

- The mean of the day of flex for cases is 60.22 days and the standard deviation is 0.77 days.
- The mean of the day of flex for fatalities is 81.4 days and the standard deviation is 0.69 days.

## Applying the day of flex indicator to the data (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia_future + geom_vline(xintercept = mean(day_of_flex_ca), color="#d33636", size=1.5, linetype="dashed") + labs(title="Day of flex indicator for cases (60 days after)"))
```

## Applying the day of flex indicator to the data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia_future + geom_vline(xintercept = mean(day_of_flex_fa), color="#1b95ff", size=1.5, linetype="dashed") + labs(title="Day of flex indicator for fatalities (81 days after)"))
```

# Confirming Results with entire dataset

## Entire data (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
cases_plot_Australia_confirm <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    geom_vline(xintercept = mean(day_of_flex_ca), color="#d33636", size=1.5, linetype="dashed") +
    labs(title = "Entire Cases in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia_confirm <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=fatalities_function_australia, color="#1b95ff", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 10)) +
    geom_vline(xintercept = mean(day_of_flex_fa), color="#1b95ff", size=1.5, linetype="dashed") + 
    labs(title = "Entire Fatalities in Australia from 6/2/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(cases_plot_Australia_confirm)
```

## Entire data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(fatalities_plot_Australia_confirm)
```

# Interesting Anomaly with United States Data

## All United States Data (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
# US data
us <- data %>% filter(location == "United States")
us <- subset(us, select = c(date, total_cases, total_deaths))
us <- us %>% mutate(days = 1:n())
us <- us %>% mutate(year = substr(date, 1, 4))
us_data_cases <- ggplot(data=us, aes(x=days, y=total_cases, color=year)) + 
  geom_point() + 
  labs(title="Total Cases in the United States from 1/22/2020 to 11/14/2022 (1,028 days)", color="Year", x="Days", y="Total Cases")
us_data_deaths <- ggplot(data=us, aes(x=days, y=total_deaths, color=year)) + 
  geom_point() + 
  labs(title="Total Deaths in the United States from 1/22/2020 to 11/14/2022 (1,028 days)", color="Year", x="Days", y="Total Deaths")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(us_data_cases)
```

## All United States Data (Fatalities)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(us_data_deaths)
```

## Conclusion

- The day of flex indicator is a good indicator to predict the time of the peak of the pandemic (for most countries).
- The gauss error function is a good for countries like...
  - China, Italy and Australia
- The gauss error function is not a good fit for countries like...
  - United States
  - Any other countries who may have different covid restrictions
- MonteCarlo simulations are neccesary due to the uncertainty of the data.


## Works Cited
- Link to the study:
  - I. Ciufolini, A. Paolozzi, Prediction of the time evolution of the Covid-19 Pandemic in Italy by a Gauss
Error Function and Monte Carlo simulations. Submitted to BioRxiv on 03.26.2020 and transferred on
03.27.2020 to MedRxiv. https://doi.org/10.1101/2020.03.27.20045104
- Link to the covid data used:
  - https://covid19.who.int/
- Source to find the second derivative of fitted function:
  - https://proofwiki.org/wiki/Derivative_of_Error_Function